---
output: github_document
always_allow_html: true
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  message = FALSE,
  warning = FALSE,
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# The portrobbo package :anchor:

The functions within this package can be used to speed up work and shorten code. It has originally been made to support a portfolio and the assignments within it but the functions are short and to the point, making them useful for various situations.

## Installation

The development version of portrobbo can be installed from [GitHub](https://github.com/) by using:

``` r
# install.packages("devtools")
devtools::install_github("mirthhe/portrobbo")
```

# Functions

The package currently contains the following functions:
<ul>
<li>uniq_val</li>
<li>save_csv_rds</li>
<li>graph_jitter</li>
<li>protein visualizations:<ul>
<li>protein_cartoon</li>
<li>protein_sphere</li>
<li>protein_stick</li>
<li>protein_visual</li></ul></ul>

To show the use of the package we first load the needed libraries.

```{r setup}
library(portrobbo)
library(ggplot2)
library(dplyr)
```

The package also contains the following data objects:

[COVID19](https://github.com/mirthhe/portrobbo/blob/main/data/covid19.rda)
<ul>
<li>A dataframe containing COVID19 cases and deaths in multiple European countries</li>
<li>Data came from the [ECDC](https://www.ecdc.europa.eu/en/publications-data/download-todays-data-geographic-distribution-covid-19-cases-worldwide)</li>
<li>Imported using [this code](https://github.com/mirthhe/portrobbo/blob/main/data-raw/covid19.R)</li>
<li>Used to show the functionality of uniq_val(), save_csv_rds() and graph_jitter()</li>
</ul>

[Kinesin](https://github.com/mirthhe/portrobbo/blob/main/data/kinesin.rda)
<ul>
<li>R object containing the PDBx file 3COB</li>
<li>[Data](https://github.com/mirthhe/portrobbo/blob/main/data-raw/3cob.pdb) came from the [RCSB PDB](https://www.rcsb.org/structure/3COB)</li>
<li>Imported using [this code](https://github.com/mirthhe/portrobbo/blob/main/data-raw/kinesin.R)</li>
<li>Used to show the functionality of the protein visualising</li>
</ul>

## uniq_val

The first function focuses on the data inspection. When working with a big dataset it can be used to check the amount of unique values within a column. For example, the dataset covid19 is all about the COVID19 cases and related deaths. The dataset has 28729 observations. Since there are multiple observations per country it is important to know how many countries are looked at. This is where uniq_val() comes in. The only input for the function is the column of which the amount of unique values should be checked. For our example it would be used as follows.

```{r uniq_val check, eval=FALSE}
# The column with the names of the countries is countriesAndTerritories
uniq_val(covid19$countriesAndTerritories)
```

The 28729 observations are divided over 30 different countries. Good to know! There are other columns, also indicating a country or region. uniq_val() can also be used to check if those other columns show the same amount of unique values, by doing the following.

```{r uniq_val compare, eval=FALSE}
# Another column indicating the country is geoId
uniq_val(covid19$countriesAndTerritories) == uniq_val(covid19$geoId)
```

## save_csv_rds

After inspecting the data you might have made a few changes in the R object, like making it tidy or extracting data that is the most useful as shown in the following code. 

```{r data extraction}
# Let's say we only want the data of summer 2020 for the Germany and France
covid19_2020 <- covid19 %>% filter(countriesAndTerritories == 
                                     c("Germany", "France"), 
                                   year == 2020, month == c(6:8))
covid19_2020$dateRep <- as.Date(covid19_2020$dateRep, "%d/%m/%y")
```

It is important to save your changes for later reference. A good way to do that is by using the save_csv_rds() function. This saves your dataframe as both a .csv and .rds file and uses the dataframe and an outputname as inputs. To use it on our covid19 dataset we would do the following.

```{r save_csv_rds, eval=FALSE}
# In this case we want the outputname to be the same as the dataframe name
save_csv_rds(covid19_2020, "covid19_2020")
```

## graph_jitter

With the data inspected, extracted and saved, graphs can be made. The ggplot2 package contains all the necessary functions to make a graph. It's functions have a lot of different input options for a lot of different graphs. This means the code for a simple graph can become quite long. This is where graph_jitter() comes in. The ggplot2 function position_jitter() uses multiple inputs to create a slight jitter to the datapoints. This is useful when data overlap makes the graph hard to read. The graph_jitter() function simplifies this function by only needing one input which will work for both the height and width of the jitter. It also sets the seed to keep jitter consistent throughout the analysis. In the following code chunk the use of graph_jitter() is shown.

```{r graph functions}
covid19_2020 %>% ggplot(aes(x = dateRep, y = deaths)) +
  geom_line(aes(group=countriesAndTerritories, color = countriesAndTerritories)) +
  geom_point(aes(color = countriesAndTerritories), position = graph_jitter(0.3)) +
  labs(title = "COVID19 in summer of 2020",
       x = "Date",
       y = "COVID19 related deaths") +
  theme_minimal()
```

<i>Figure 1. Scatterplot showing the COVID19 related deaths in summer of 2020 in Germany and France.</i>

## Protein visualization fucntions

The last three functions use the r3dmol package to create 3D interactive visualizations of proteins. Each function shows a different style of visualization. The input is a PDB file turned into an R object using the readLines() function. For this example we will be looking at my favorite protein, kinesin, which is included in the package. The PDBx file for kinesin was taken from [3COB](https://www.rcsb.org/structure/3COB) in the RCSB database. The colors are modeled after the PyMol default colors.

### protein_cartoon

The first visualization is the cartoon style. It is a simplified version where sidechains are ignored. Only the backbone is shown with the alpha helices as coils and the beta sheets as arrows.

```{r protein_cartoon}
protein_cartoon(kinesin)
```

<i>Figure 2. Kinesin as described by the 3COB PDBx file from wwPDB. Visualised in cartoon style using the r3dmol package and PyMOL color scheme. Using a click and drag method the image can be turned and by scrolling the image can be zoomed in and out.</i>

### protein_sphere

Spheres, unlike cartoon, shows the atoms. Each atom is shown as a sphere. The size of each sphere is decided by the atoms electron orbitals. This can be used to show packing and spheric hindrance but doesn't show the interior. The colors indicate different atoms, in the default settings carbon atoms are green, nitrogen is blue, oxygen is red, etc.

```{r protein_sphere}
protein_sphere(kinesin)
```

<i>Figure 3. Kinesin as described by the 3COB PDBx file from wwPDB. Visualised in sphere style using the r3dmol package and PyMOL color scheme. Using a click and drag method the image can be turned and by scrolling the image can be zoomed in and out.</i>

### protein_stick

The last visual style will be sticks, this shows the chemical structure. The colors still indicate the same atoms as they did for the sphere style.

```{r protein_stick}
protein_stick(kinesin)
```

<i> Figure 4. Kinesin as described by the 3COB PDBx file from wwPDB. Visualised in stick style using the r3dmol package and PyMOL color scheme. Using a click and drag method the image can be turned and by scrolling the image can be zoomed in and out.</i>

### protein_visual

A big downside of these visualizations is how none of them show a complete picture. Cartoon style doesn't show the atoms, sphere styles doesn't show the chemical structure and stick style doesn't show the sferic hindrance. Because of this, one more function was made, to combine the stick and sphere visuals in a way that clearly shows both of them. To do this they were combined in one function with the stick visual on full opacity and the sphere visual on an opacity of 0.7. The clarity of the spherical style could be better and the overlap definitely has a negative effect on the readability, but the combination still gives a good overview of the protein's 3D structure.

```{r protein_visual}
protein_visual(kinesin)
```

<i> Figure 5. Figure 9.7 Kinesin as described by the 3COB PDBx file from wwPDB. Visualised in stick and spherical style using the r3dmol package and PyMOL color scheme. Using a click and drag method the image can be turned and by scrolling the image can be zoomed in and out.</i>
